---
title: "Assignment 3"
output: html_notebook
Dominique Schreiner
---

NOTES: 
lookup tables needed: 
[X]subject
[X] certificate

Subject is categorical and Topic has ONLY 1 subject 
Only Certifications can be NULL everything is needs to be NOT NULL

From dev.mysql.com

"CASCADE : Delete or update the row from the parent table and automatically delete or update the matching rows in the child table. ...
SET NULL : Delete or update the row from the parent table and set the foreign key column or columns in the child table to NULL ."
```{r}
library(RSQLite)



fpath = "/Users/domschrein/Desktop/SQLite/"
dbfile = "commiteeDB.sqlite"

dbcon <- dbConnect(RSQLite::SQLite(), paste0(fpath,dbfile))
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Author

```


```{sql connection=dbcon}
CREATE TABLE Author(
  aid INTEGER NOT NULL,
  name varchar(255) NOT NULL,
  email varchar(255) NOT NULL,
  PRIMARY KEY(aid)
)
```

```{sql connection=dbcon}
SELECT * FROM Author;

```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Certifications
```

Check to see if the column on pertains the data in the list below
Create a lookup table for the Certification Attribute in Author 
```{sql connection=dbcon}
CREATE TABLE Certifications(
  cid INTEGER NOT NULL,
  cert_title VARCHAR(50) NOT NULL,
  CHECK (cert_title IN('PMP','CBAP','CSM','CSTE', 'CAP'))
  PRIMARY KEY(cid)
)
```

Using a join statement to link the columns that hold certiticate title from Certificate table to the Author's name in Author Table 
Author can have more than one certificate, and certificates can many authors (many to many)
```{sql connection=dbcon}
SELECT Author.name, Certifications.cert_title
FROM Certifications
JOIN Author ON Author.name = Certifications.cert_title;

```
```{sql connection=dbcon}
DROP TABLE IF EXISTS Topic 
```

Topic has exactly one subject, therefore I will use the UNIQUE constraint to accommodate this condition, I am assuming that the topic id can have the same length as another, however since the topic has exactly one subject, a same title would be unlikely, but not impossible. However, for this assignment I'm assuming it would differ.
```{sql connection=dbcon}
CREATE TABLE Topic(
  tid INTEGER NOT NULL,
  title varchar(255) NOT NULL,
  lengthInHrs INTEGER NOT NULL,
  subject varchar(255) NOT NULL,
  aid INTEGER NOT NULL,
  PRIMARY KEY (tid),
  FOREIGN KEY(aid) REFERENCES Author(aid) ON DELETE SET NULL,
  FOREIGN KEY(subject) REFERENCES Subject(sub_id) ON DELETE CASCADE,
  CONSTRAINT subject_constraint UNIQUE(subject, title)
)
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Subject
```

creating a subject lookup table with foreign keys as the referential element linking to Topic 
```{sql connection=dbcon}
CREATE TABLE Subject(
  sub_id INTEGER NOT NULL,
  sub_name varchar(255) NOT NULL, 
  PRIMARY KEY (sub_id),
  CONSTRAINT sub_unique UNIQUE(sub_id, sub_name),
  CHECK (sub_name in ('ML', 'SQL', 'R', 'Java'))  
)

```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Course 
```

1 CourseTopic per Course relation 
```{sql connection=dbcon}
CREATE TABLE Course(
  course_number varchar(255) NOT NULL,
  title varchar(255) NOT NULL,
  lenghtInDays INTEGER NOT NULL,
  CONSTRAINT course_constraint UNIQUE(title),
  PRIMARY KEY(course_number)
) 
  
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS CourseTopic 
```

linking table to display many to many relationship of Course and Topic
```{sql connection=dbcon}
CREATE TABLE CourseTopic(
  course_code varchar(255) NOT NULL,
  topic_code INTEGER NOT NULL,
  FOREIGN KEY(course_code) REFERENCES Course(course_number) ON DELETE CASCADE,
  FOREIGN KEY(topic_code) REFERENCES Topic(tid) ON DELETE CASCADE
  
)

```

```{r}
dbDisconnect(dbcon)
```

