---
title: "Assignment 9: Query XML "
output: html_notebook
---


```{r}
#install.packages("XML")
library(XML)
```


parsing XML file
```{r}
path <- "/users/domschrein/desktop/"
xmlFile <- "Books-v3.xml"
fp <- paste0(path,xmlFile)


xmlObj <- xmlParse(fp)
```



QUERY 1:
What are the titles of all books written by "Galos, Mike"? List the titles and the prices.
```{r}
xpathEx <- "/catalog/book[author='Galos, Mike']/title|/catalog/book[author='Galos, Mike']/price"

title_price <- xpathSApply(xmlObj, xpathEx, xmlValue)

title_price
```



QUERY 2:
In what genre is the book with id "bk109".
```{r}

xpathEx <- '/catalog/book[@id ="bk109"]/genre/text()'
bk109 <- xpathSApply(xmlObj, xpathEx, xmlValue)
bk109
```


QUERY 3:
What are the titles of the books that sell for the lowest price? (Use the min aggregation function)
..[../price=min(//price)]
```{r}

xpathEx <- '/catalog/book[price= min(//price)]/title'

min_price <- xpathSApply(xmlObj, xpathEx, xmlValue)
min_price
```



QUERY 4: 
List the title and prices of all books in the "Computer" genre.
```{r}
xpathEx <- '/catalog/book[genre ="Computer"]/title|/catalog/book[genre ="Computer"]/price'
genre <- xpathSApply(xmlObj, xpathEx, xmlValue)
genre
```


QUERY 5: 
What is the number of all books that cost less than $5?
```{r}
xpathEx <- 'count(/catalog/book[price>5]/title/text())'
cheap_books <- xpathSApply(xmlObj, xpathEx, xmlValue)
cheap_books
```





