---
title: "CS 5200 Assignment 6"
output: html_notebook
creator: Dominique Schreiner

---

Triggers are automated callback functions, that occur when events occur on the secified table or if defined by row using FOR EACH ROW. Such events included INSERT, DELETE, OR UPDATE. 

Audit Triggers are used to log (to track) these type of functions mentioned above. In my trigger I will be creating an audit trail for every of inventory for the book of the  Author table. 

```{r}
library(RSQLite)

fpath = "/Users/domschrein/Desktop/CS5200/"
dbfile = "commiteeDB.sqlite"

dbcon <- dbConnect(RSQLite::SQLite(), paste0(fpath,dbfile))
```


```{sql connection=dbcon}
DROP TABLE IF EXISTS "Author"
```

Added a few columns to use for my Audit Trigger. "publishinghouse",
"titlespublished", "inventory" to see if I can log the inventory and type of event occurring while maintaing the integrity of the other data. 

This is the table specified and the columns involved are "aid", "inventory" 
```{sql connection=dbcon}
CREATE TABLE "Author"(
  "aid" INTEGER PRIMARY KEY NOT NULL,
  "name" varchar(255) NOT NULL,
  "email" varchar(255) NOT NULL,
  "title" VARCHAR(255) NOT NULL,
  "publishinghouse" VARCHAR(255) NOT NULL,
  "titlespublished" INTEGER NOT NULL,
  "inventory" INTEGER NOT NULL
  
);
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS "Author_Audits"
```

Here we are creating the Audit table to track all the  "inventory" log transactions.
"aid" will be this table's ID as it comes from Author.Id.
"TRIGGER_AuthorAudits" will reflect the type of trigger 
"entry_date" will keep the timestamp when the record will be created. 
```{sql connection=dbcon}
CREATE TABLE "Author_Audits"(
  "aid" INTEGER NOT NULL,
  "TRIGGER_AuthorAudits" NOT NULL,
  "entry_date" TEXT NOT NULL
);

```

Here I am creating an insert trigger on Author Table, which will log the insertion of a record
```{sql connection=dbcon}
CREATE TRIGGER insert_log AFTER INSERT

ON "Author"

BEGIN

   INSERT INTO "Author_Audits"("aid", "entry_date", "TRIGGER_AuthorAudits") VALUES (new.aid, datetime('now'), 'INSERT');

END;
```


Here I am creating an update trigger on Author Table, which will log the update of a record
```{sql connection=dbcon}
CREATE TRIGGER update_log AFTER UPDATE

ON "Author"

BEGIN

  INSERT INTO "Author_Audits"("aid", "entry_date", "TRIGGER_AuthorAudits") VALUES (new.aid, datetime('now'), 'UPDATE');

END;
```


Here I am creating a delete trigger on Author Table, which will log the deletion of a record
```{sql connection=dbcon}
CREATE TRIGGER delete_log AFTER DELETE

ON "Author"

BEGIN

   INSERT INTO "Author_Audits"("aid", "entry_date", "TRIGGER_AuthorAudits") VALUES (new.aid, datetime('now'), 'DELETE');

END;

```

TESTING THE TRIGGER
This is where I will insert a record into Author table which should result in creating an insert log record in Author_Audits table. 
```{sql connection=dbcon}
INSERT INTO "Author" (
  "aid",
  "name",
  "email",
  "title",
  "publishinghouse",
  "titlespublished",
  "inventory"
  )

VALUES (1184, 'Jones James', "JJs@JJwrites.com", 'Ph.D', "Morton", 3, 489)
```

```{sql connection=dbcon}
SELECT * FROM "AUTHOR"
```


```{sql connection=dbcon}
SELECT * FROM "Author_Audits";
```

Similarly we execute UPDATE or DELETE operations on the Author_Audits table, log record also be updated, using the code below I will test the UPDATE event:
```{sql connection=dbcon}
UPDATE "Author" SET "inventory"= 488 WHERE "aid"= 1184;
```

Here we can see the UPDATE Sucessfully implemented along with the insert. 
Note the "aid" must match from the insert to make changes upon the record when testing the UPDATE fucntion.
```{sql connection=dbcon}
SELECT * FROM "Author_Audits";
```


Here we can see all the triggers from sqlite_master table, using following code:
```{sql connection=dbcon}
SELECT name FROM sqlite_master

WHERE type = 'trigger' AND tbl_name = "Author";

```

One can also delete the triggers using DROP command which is as follows:
```{sql connection=dbcon}
DROP TRIGGER 'insert_log'; 
```

We can also see all the triggers currently logged on Author table, then use AND clause with table name as follows:
```{sql connection=dbcon}
SELECT name FROM sqlite_master

WHERE type = 'trigger' AND tbl_name = "Author";
```

```{r}
dbDisconnect(dbcon)
```

